Ornaments as indicators of social changes in northeastern Taiwan before and after the European colonial period

### 2020-04-19 analysis/paper/paper.Rmd
reorganize the conclusion and shorten the future work

Ornaments (Figure \@ref(fig:ornaments-image)) were found in a variety of archaeological contexts including post-holes area, burials, and middens. For the ornament categories, we follow the well established topology based on raw materials and shapes for the region of northeastern Taiwan [@Chen2007; @NMTH2005]. This study focuses on `r sam_post_hole` ornaments from 40 sampling squares located at the dwelling place of Kiwulan, indicated by aligned post-holes with _in-situ_ posts (Figure \@ref(fig:KWL-sam-area)). Occupation floors were not identified during excavation. We choose these units because they were stratigraphically intact and undisturbed by modern construction activity, compared to excavation squares on the periphery of the site. There are 35 burials in the sampling area, one third of the total number of burials at Kiwulan. Ornaments are commonly used as grave goods in burials, with the total number of ornaments in burials including 3,173. Besides, there are 27 ornaments found in midden contexts. The obvious difference in number in burials and midden contexts indicates the property of ornaments as prestige goods. The high number in burials is due to the presence of bead strands or patterned bands of beads, which sometimes contains thousands of beads in an individual burial [@Chen2007]. The unequal distribution of beads between burials was interpreted as the evidence for hierarchy by @Cheng2008. However, @Hsieh2012 suggests a relatively egalitarian form from the comparative analysis of frequency and proportion of burial goods. To have a comprehensive understanding, we focus on ornaments from the dwelling place contexts (Figure \@ref(fig:plot-all-ornaments-count), Table \@ref(tab:tidy-data-for-table)) because these give us the greatest spatial and temporal representation across the three time periods, and so are most informative of social inequality as indicated by uneven distributions of ornaments. The burials are excluded because most burials from the sampling area date to the European period (n = 21), limiting the usefulness of comparisons between the periods. Ornaments could be potential heirlooms inherited over multiple generations. However, there is no continuous increase over time that suggests accumulation and discard of ornaments is not constant. Thus, we assume that changes in the abundance of ornaments reflect changes in the amount in archaeological record. Houses and burials may also be a useful source of evidence to understand population size but a proper treatment of those is beyond the scope of the paper.

@@

An indirect colonial influence may be indicated at Kiwulan by the greater diversity of ornament types and materials during the European period. Yilan was involved in complex trading networks both on a regional scale with other indigenous groups and Chinese merchants, and at a global scale with Europeans, including the Dutch and the Spanish. The ornaments have multiple origins, including Southeast Asia and China, and were first introduced into northeastern Taiwan by Chinese merchants before 17th century. Later, trade activities became more frequent and intense in the 17th century due to European activities. The greater diversity and quantity of ornaments likely resulted from participation in large scale exchange networks that stimulated the circulation of different ornament classes. The frequency of overall ornaments and each subtype declines significantly after European influence fades during the Chinese period in the early 19th century. This may be due to a smaller scale of trading networks, the overall decline of indigenous populations in Yilan, or the adoption of Han Chinese practices. The decline of the population at Kiwulan may be related to the movement of many indigenous people southwards to Hualien due to the increasing Han Chinese immigrants who took over their lands at the end of the 18th century [@Chen2007].

@@

We are still far from understanding the full variety of colonial impacts on peripheral indigenous communities. The origin of ornaments and how their sources changed over time would give more information to construct a clear picture of complex trade networks during this periods. By focusing on the distribution patterns in a settlement site, the Kiwulan ornaments suggest that foreign ornaments can be a proxy to detect indirect colonial influence on local indigenous populations. Ornaments give insights into the amplification of social inequality stimulated by European colonization. It also shows the agency of indigenous people to incorporate ornaments into their social system and use them in their daily lives to display or intensify status differences. Future work could extend this approach to sourcing the ornaments directly with geochemical methods, and studies of other trade goods such as ceramics. We have introduced here the corporate/network model for understanding the dynamics of social inequality at Kiwulan, and future tests of this should include analysis of pottery production and standardization, and mortuary practices.

>>>>

Ornaments (Figure \@ref(fig:ornaments-image)) were found in a variety of archaeological contexts including post-holes area, burials, and middens. For the ornament categories, we follow the well established topology based on raw materials and shapes for the region of northeastern Taiwan [@Chen2007; @NMTH2005]. This study focuses on `r sam_post_hole` ornaments from 40 sampling squares located at the dwelling place of Kiwulan, indicated by aligned post-holes with _in-situ_ posts (Figure \@ref(fig:KWL-sam-area)). Occupation floors were not identified during excavation. We choose these units because they were stratigraphically intact and undisturbed by modern construction activity, compared to excavation squares on the periphery of the site. There are 35 burials in the sampling area, one third of the total number of burials at Kiwulan. Ornaments are commonly used as grave goods in burials, with the total number of ornaments in burials including 3,173. Besides, there are 27 ornaments found in midden contexts. The obvious difference in number in burials and midden contexts indicates the property of ornaments as prestige goods. The high number in burials is due to the presence of bead strands or patterned bands of beads, which sometimes contains thousands of beads in an individual burial [@Chen2007]. The unequal distribution of beads between burials was interpreted as the evidence for hierarchy by @Cheng2008. However, @Hsieh2012 suggests a relatively egalitarian form from the comparative analysis of frequency and proportion of burial goods. To have a comprehensive understanding, we focus on ornaments from the dwelling place contexts (Figure \@ref(fig:plot-all-ornaments-count), Table \@ref(tab:tidy-data-for-table)) because these give us the greatest spatial and temporal representation across the three time periods, and so are most informative of social inequality as indicated by uneven distributions of ornaments. The burials are excluded because most burials from the sampling area date to the European period (n = 21), limiting the usefulness of comparisons between the periods. Ornaments could be potential heirlooms inherited over multiple generations. However, there is no continuous increase over time that suggests accumulation and discard of ornaments is not constant. Thus, we assume that changes in the abundance of ornaments reflect changes in the amount in archaeological record.

@@

An indirect colonial influence may be indicated at Kiwulan by the greater diversity of ornament types and materials during the European period. Yilan was involved in complex trading networks both on a regional scale with other indigenous groups and Chinese merchants, and at a global scale with Europeans, including the Dutch and the Spanish. The ornaments have multiple origins, including Southeast Asia and China, and were first introduced into northeastern Taiwan by Chinese merchants before 17th century. Later, trade activities became more frequent and intense in the 17th century due to European activities. The greater diversity and quantity of ornaments likely resulted from participation in large scale exchange networks that stimulated the circulation of different ornament classes. The frequency of overall ornaments and each subtype declines significantly after European influence fades during the Chinese period in the early 19th century. This may be due to a smaller scale of trading networks, the overall decline of indigenous populations in Yilan, or the adoption of Han Chinese practices. The decline of the population at Kiwulan may be related to the movement of many indigenous people southwards to Hualien due to the increasing Han Chinese immigrants who took over their lands at the end of the 18th century [@Chen2007]. Houses and burials may also be a useful source of evidence to understand population size but a proper treatment of those is beyond the scope of the paper.

@@

We are still far from understanding the full variety of colonial impacts on peripheral indigenous communities. The origin of ornaments and how their sources changed over time would give more information to construct a clear picture of complex trade networks during this periods. We have introduced here the corporate/network model for understanding the dynamics of social inequality at Kiwulan, and future analysis of pottery production and standardization, and mortuary practices would provide more information. By focusing on the distribution patterns in a settlement site, the Kiwulan ornaments suggest that foreign ornaments can be a proxy to detect indirect colonial influence on local indigenous populations. Ornaments give insights into the amplification of social inequality stimulated by European colonization. It also shows the agency of indigenous people to incorporate ornaments into their social system and use them in their daily lives to display or intensify status differences.
